\begin{enumerate}
    \item Ageing Up
    \begin{enumerate}
        \item The game has a button which “ages up” the character
        \item As part of this process, a random event will occur (eg, family member dies, your partner wins something, you are pregnant)
        \item Some events will happen at set times (eg. finishing school)
    \end{enumerate}
    \item Relationships
    \begin{enumerate}
        \item Having a relationship with parents
        \item Having partners
        \item Marrying partners
        \item Divorcing partners
        \item Being able to pick the characters sexuality
        \item Having kids with your partner (adopted/biological)
    \end{enumerate}
    \item Crime
    \begin{enumerate}
        \item Ability to commit crime - this will be a button which will randomly generate a crime, then deduct a set amount from the lifescore.
    \end{enumerate}
    \item Employment
    \begin{enumerate}
        \item Get jobs
        \item Quit jobs
        \item The longer you stay in a job, the more \verb|jobPoints| a character will get each year. 
    \end{enumerate}
    \item Character Picker
    \begin{enumerate}
        \item At the start of each life, the user will be able to pick the character they want to play the game as. This will generate a brief backstory to the character which will crop up throughout the game (deaths of family members, key events in siblings' lives)
    \end{enumerate}
    \item Education
    \begin{enumerate}
        \item There will be a random education element to the game which will allow players to progress through school (which will have randomly generated names). Every time they complete a course or school, they get \verb|educationPoints|. 
    \end{enumerate}
    \item Saving
    \begin{enumerate}
        \item The game should be able to write the data out to a file(s) which can be loaded in during runtime to resume the game. These do not need to be readable by the user.
        \item There should be a way for the user to export the life from the game at the end of the game to a non-editable file format (eg, PDF).
    \end{enumerate}
    \item Scores
    \begin{enumerate}
        \item There should be a number of scores which the actions of the player can affect throughout the game.
    \end{enumerate}
\end{enumerate}