\chapter{Introduction}

\section{Introduction to the Project}
Traffic lights are fundamental to the smooth running of the road networks. Without them vehicle drivers would be placed in potentially dangerous situations multiple times per journey, where they approach junctions with multiple roads; not to mention the risk which pedestrians would take every time they cross the road. Traffic lights (amongst other electronically controlled traffic management systems) provide some order to what would otherwise be uncontrolled chaos.\newline
I chose those this project with the aim of understanding how a traffic light control system works as well as understanding how microcontrollers work, interfacing with their inputs and outputs. Whilst the unknown of how microcontrollers work added to the complexity, the comfort of using traffic lights daily makes this feel like its a manageable task to complete in the time allocated. 

\section{Definition of Traffic Lights}
Everyone knows what traffic lights do, but for this project I feel it is deeply important to have an understanding of the definition of a traffic lights. \newline
Collins English Dictionary\footnote{https://www.collinsdictionary.com/dictionary/english/traffic-light} defines 'Traffic Lights' as:
\blockQuote{Traffic lights are sets of red, amber, and green lights at the places where roads meet. They control the traffic by signalling when vehicles have to stop and when they can go. Traffic lights can also be referred to as a traffic light.}